\documentclass[11pt,a4paper]{article}

\input{common/dct_preamble}

\title{
    \bfseries On the Necessity of Codimension-2 Reduction in Resolving Spacetime Singularities
    }
    
\author{
    Marek Hubka\, \orcidlink{0009-0003-2476-9017}
    \thanks{
        Independent Researcher, Czech Republic. \\
        Website: \href{http://tidesofuncertainty.com}{tidesofuncertainty.com}.
        Email: \href{mailto:marek@tidesofuncertainty.com}{marek@tidesofuncertainty.com}.
    }
}
\date{\today}

\begin{document}
\maketitle
\vspace{-2ex}

\begin{abstract}
This article motivates the foundational axiom of Dimensional Collapse Theory
(DCT/QG): that when curvature reaches a universal threshold, spacetime undergoes
a local, codimension-2 ``snap'' which removes a null 2-plane and leaves behind
a spacelike \((D-2)\)-dimensional surface---the Ledger~\(\LL\). The aim is not
to derive this rule from deeper principles, but to show that it is a highly
constrained and natural choice once we impose a small set of physical demands:
locality, covariance, causality, boost invariance, and compatibility with
area-based thermodynamics. We review the singularity problem as a geometric
crisis driven by the Raychaudhuri equation, enumerate possible local geometric
responses at a curvature threshold, and argue that a codimension-1 or higher
codimension reduction fails to provide a stable, thermodynamic boundary. In
contrast, deleting exactly the two null directions normal to a collapsing
cross-section yields a unique ``Goldilocks'' option: it halts catastrophic
focusing, preserves causal structure, creates a natural area carrier for
entropy, and sets the stage for the detailed constructions in Papers~I--IV of
the DCTQG series.
\end{abstract}

\tableofcontents

\section{Singularities as a geometric crisis}

The classical singularity theorems show that, under broad conditions, General
Relativity predicts geodesic incompleteness and curvature blow-ups
\cite{HawkingEllis1973}. The underlying mechanism is not mysterious: it is the
relentless focusing of geodesic congruences.

For a null congruence with tangent \(k^{a}\), expansion \(\theta\), and affine
parameter \(\lambda\), the Raychaudhuri equation in \(D\) spacetime dimensions
is \cite{Raychaudhuri1955,Poisson2004}
\begin{equation}
\frac{\dd\theta}{\dd\lambda}
=
-\frac{1}{D-2}\,\theta^{2}
- \sigma_{ab}\sigma^{ab}
- R_{ab}k^{a}k^{b},
\qquad
\frac{\dd \ln \Acal}{\dd\lambda} = \theta,
\label{eq:raychaudhuri}
\end{equation}
where \(\sigma_{ab}\) is the shear and \(\Acal(\lambda)\) is the cross-sectional
area of an infinitesimal bundle. Under the null energy condition (NEC),
\(R_{ab}k^{a}k^{b}\ge 0\), so a negative \(\theta\) typically runs away to
\(-\infty\) in finite affine parameter. The area \(\Acal\) shrinks to zero and
curvature invariants diverge.

From the perspective of effective field theory, this signals a \emph{geometric
crisis}: the classical description runs out before we reach the regime we would
like to probe. A satisfactory resolution should obey at least four conditions:
\begin{enumerate}[label=(\alph*)]
\item \textbf{Locality and covariance:} the rule must be formulated locally in
terms of tensorial quantities, without reference to global coordinates or
boundary conditions.

\item \textbf{Causality:} the resolution must not introduce superluminal
signaling or retro-causal behavior.

\item \textbf{Thermodynamic compatibility:} in the presence of horizons, any
resolution should respect the area-entropy connection
\cite{Bekenstein1973,Hawking1975,Wald1993} and the generalized second law
\cite{Bousso1999}.

\item \textbf{Mildness:} the modification should be as small and surgical as
possible, leaving ordinary GR+QFT intact where curvature is subcritical.
\end{enumerate}

Dimensional Collapse Theory starts from the idea that the breakdown occurs
because we are trying to continue the geometry \emph{through} a regime in which
the classical dimensionality of spacetime is no longer the correct
description. The central proposal is that spacetime carries a simple, local
``click rule'': when a clean, dimensionless invariant
\(\Ical[g]\) crosses a universal threshold \(\Icrit\)
\cite{DCTQG01}, the manifold undergoes a small, discrete change in its
effective dimensionality.

\section{Local curvature triggers and click rules}

The first step is to ensure that the trigger is geometric and local. In DCT we
use a curvature scalar of the schematic form
\begin{equation}
\Ical \;\equiv\; K L^{4},
\qquad
K = R_{ABCD}R^{ABCD},
\end{equation}
where \(L\) is a macroscopic length scale set by the environment
\cite{DCTQG01}. The details of the choice and the derivation of the universal
critical value \(\Icrit\) are the subject of Paper~I. For present purposes we
only need three qualitative properties:
\begin{enumerate}[label=(\roman*)]
\item \(\Ical\) is scalar and local.
\item \(\Ical\) is dimensionless and insensitive to trivial rescalings.
\item There exists a universal critical value \(\Icrit\) such that, in
astrophysical black holes, \(\Ical\to\Icrit\) just inside the event horizon.
\end{enumerate}

The click rule is:
\begin{equation}
\Ical \;=\; \Icrit
\quad\Longrightarrow\quad
\text{spacetime performs a local \emph{snap}}.
\end{equation}
The rest of the DCTQG series works out ``what the snap does'' in detail. This
paper focuses on one key aspect: \emph{how many dimensions are removed} and
\emph{why}.

\section{A menu of geometric responses}

Suppose that in some small neighborhood the curvature grows and
\(\Ical\to\Icrit\). What can the geometry do, locally and covariantly, to avoid
the catastrophic focusing implied by Eq.~\eqref{eq:raychaudhuri}?

At an abstract level, there are three broad classes of response:
\begin{enumerate}[label=(\alph*)]
\item \textbf{Modify the dynamics:} change the right-hand side of the
Raychaudhuri equation by violating the energy conditions, introducing new
repulsive forces, or changing the Einstein equations themselves.

\item \textbf{Impose nonlocal structure:} introduce global matching conditions,
wormholes, or other topological identifications which redirect the geodesics
elsewhere.

\item \textbf{Modify the manifold structure:} change the dimensionality of the
effective spacetime in the region where the crisis occurs.
\end{enumerate}

Options (a) and (b) are widely explored in the literature. DCT instead pursues
option (c): keep the field equations and local energy conditions intact as long
as possible, and let spacetime itself take a small, discrete step in
dimensionality when the click rule fires. The question then becomes sharp:
\emph{what codimension \(\Delta D\) is admissible?}

\section{Why codimension one is not enough}

A seemingly mild modification is to remove a single null direction, i.e.\ to
replace a \(D\)-dimensional region by a \((D-1)\)-dimensional null hypersurface
when \(\Ical=\Icrit\). This effectively projects the dynamics onto a lightfront
and might appear to ``cap off'' the focusing.

However, codimension-one reduction fails on two fronts:

\paragraph{(1) Focusing persists within the null hypersurface.}
A null hypersurface still admits null congruences tangent to it, with their own
expansions and shears. The Raychaudhuri equation continues to drive
\(\Acal\to 0\) \emph{within} the reduced geometry. One has simply moved the
problem from the bulk to the boundary, not removed it.

\paragraph{(2) Null surfaces do not carry a robust area entropy.}
Bekenstein--Hawking entropy and the generalized entropy functionals used in
quantum extremal surface (QES) constructions are built on \((D-2)\)-dimensional
spacelike surfaces with well-defined area \cite{Bekenstein1973,Hawking1975,
Wald1993,EngelhardtWall2015}. The ``area'' of a patch on a null hypersurface
depends on an arbitrary choice of cross-section and affine parameter; it cannot
serve as a stable, frame-independent measure of thermodynamic entropy. Any
theory that aims to preserve the area--entropy connection in the spirit of
Jacobson's derivation of Einstein's equations \cite{Jacobson1995} must treat
null sheets as \emph{carriers} of flux, not as the fundamental storage surface
for entropy.

For DCT, which explicitly uses area quanta on a codimension-2 Ledger to
account for information payloads \cite{DCTQG03}, a codimension-1 reduction
would be a poor foundation.

\section{Codimension-two removal as the ``Goldilocks'' option}

DCT instead adopts a codimension-two operation: a local removal of the normal
null 2-plane spanned by \((\nplus,\nminus)\). Concretely, pick null directors
\(\nplus^{A},\nminus^{A}\) satisfying
\begin{equation}
\nplus\!\cdot\nminus=-1,
\end{equation}
and define the tangential projector
\begin{equation}
P^{A}{}_{B}
=
\delta^{A}{}_{B}
+ \nplus^{A}\nminus_{B}
+ \nminus^{A}\nplus_{B}.
\end{equation}
The Null--Pair Removal (NPR) map of Paper~II \cite{DCTQG02} contracts every
index of every tensor with \(P^{A}{}_{B}\), deleting components along the normal
null 2-plane while preserving the intrinsic \((D-2)\)-dimensional geometry.

This choice satisfies the constraints listed in Sec.~1 in a remarkably tight
way:
\begin{enumerate}[label=(\alph*)]
\item \textbf{Minimality.} Removing exactly the two null directions normal to a
collapsing cross-section is the smallest possible change that directly targets
the focusing problem in Eq.~\eqref{eq:raychaudhuri}. Higher-codimension
removals would overkill the geometry and destroy more structure than necessary.

\item \textbf{Boost invariance.} A single null direction has no invariant
normalization; rescalings \(k^{a}\to \ee^{\lambda}k^{a}\) are an exact local
symmetry. The pair \((\nplus,\nminus)\), by contrast, carries an invariant
boost class under \(\mathrm{SO}(1,1)\): \(\nplus\to \ee^{\lambda}\nplus\),
\(\nminus\to \ee^{-\lambda}\nminus\). The projector \(P^{A}{}_{B}\) is built to
be invariant under this symmetry. Any local rule that singles out one null
direction over the other would explicitly break this gauge freedom; NPR does
not.

\item \textbf{Thermodynamic canvas.} The image of \(P^{A}{}_{B}\) is a
spacelike \((D-2)\)-dimensional surface \(\LL\) with intrinsic metric
\(h_{ab}\) and area \(A(\LL)\). This provides a natural ``canvas'' for
area-based entropy. In Paper~I, this surface is placed at a universal radius
inside a Schwarzschild black hole via the curvature trigger \(\Ical=\Icrit\)
\cite{DCTQG01}, and in Paper~III its area is quantized by the rules of
Transdimensional Thermodynamics (TDT) \cite{DCTQG03}.

\item \textbf{Information accounting and reversibility.} Removing a null
2-plane is not purely topological: it deletes geometric data (orientation,
shear alignment, twist) that can, in principle, be recorded. This is precisely
what the Infinity bits \(\{W,X,Y,Z\}\) of Paper~IV are designed to capture
\cite{DCTQG04}. NPR plus these bits make the snap reversible up to smooth
gauge, in the same way that recording a small set of discrete choices restores
reversibility to a coarse-grained map.
\end{enumerate}

In this sense, codimension-two reduction is a ``Goldilocks'' operation: any
less and the singularity problem persists; any more and one loses the geometric
structure needed to define a thermodynamic boundary.

\section{How the axiom organizes the DCTQG series}

The codimension-2 snap is the only genuine axiom of DCT/QG. Everything else in
the series is derived from, or calibrated against, this geometric choice.

\begin{itemize}[leftmargin=*]
\item \textbf{Part I.} Given the click rule~\(\Ical=\Icrit\), the
location of \(\LL\) in Schwarzschild and more general black-hole geometries is
computed explicitly, and the universal constant \(\T\) is derived from the
requirement that Ledger and horizon entropies remain consistent with
Bekenstein--Hawking and QES bounds \cite{DCTQG01}.

\item \textbf{Part II.} NPR is worked out in detail: the precise
form of \(P^{A}{}_{B}\), its boost invariance, and its action on fields of
arbitrary rank are specified. The Raychaudhuri equation is updated to a form
that remains finite in the presence of snaps \cite{DCTQG02}.

\item \textbf{Part III.} Transdimensional Thermodynamics is
constructed on \(\LL\): one bit of payload entropy costs a fixed quantum of
area, and the Ledger entropy grows in a staircase fashion at snaps while
remaining constant between them \cite{DCTQG03}.

\item \textbf{Part IV.} The Infinity bits \(\{W,X,Y,Z\}\) are
introduced as a minimal, boost-invariant register that records discrete
kinematic information about the removed null 2-plane, making the NPR operation
reversible up to smooth gauge and allowing a clean information-theoretic
interpretation of snaps \cite{DCTQG04}.
\end{itemize}

Seen from this vantage point, the codimension-2 axiom is analogous to the
Schr\"odinger equation in quantum mechanics: it is not derived from something
deeper within the framework, but it proves to be the unique simple postulate
from which a large, tightly constrained structure grows.

\section{Conclusion}

The dimensional-collapse axiom of DCT/QG can be stated in one line:
\emph{when a local curvature invariant \(\Ical\) reaches a universal threshold
\(\Icrit\), spacetime undergoes a codimension-2 snap that removes the normal
null 2-plane and leaves a spacelike Ledger surface \(\LL\) behind.} This
paper has argued that, once we insist on locality, covariance, causality,
boost invariance, and compatibility with area-based thermodynamics, this is not
a whimsical choice among many, but a highly constrained option.

A codimension-1 reduction fails to halt focusing and cannot carry a stable
entropy. Higher-codimension reductions destroy too much structure. Only the
codimension-2 NPR operation both disarms the singularity mechanism and
provides a natural canvas for entropy, information accounting, and reversible
reconstruction. The success of DCT/QG as a whole---from Ledger placement and
TDT to Infinity bits and echoes phenomenology---is, in this sense, the primary
evidence in favor of this simple axiom.

\bibliographystyle{unsrt}
\bibliography{bibliography/dct_refs,bibliography/external_refs}

\end{document}